---
title: "p8105_hw1_jv2629"
output: github_document
---
##Initial commit

#Problem 1
##Loading tidyverse to enable tibble and create a data frame
```{r making_first_df}
library(tidyverse)
```

##Creating data frame with random variables
```{r problem1_df}
problem1_df = tibble(
  random_samp10 = rnorm(10, mean = 0, sd = 5),
  more_than_2 = ifelse (random_samp10 > 2,"yes", "no"),
  vec_cities = c("Boston", "Miami", "NYC", "LA", "Amsterdam", "Dar es Salaam", "Kampala", "Tel Aviv", "Rome", "Stockholm"),
  vec_continent = factor(c("North America", "North America", "North America", "North America", "Europe", "Africa", "Asia", "Asia", "Europe", "Europe"))
)

help("ifelse")

problem1_df
```
The data frame looks as I would expect it to look. The conditional statement appears to work properly. The character and factor variables match up as they should.

##Taking the mean of each of the variables
```{r mean_trials}
mean(problem1_df[["random_samp10"]])
mean(problem1_df[["more_than_2"]])
mean(problem1_df[["vec_cities"]])
mean(problem1_df[["vec_continent"]])
```
Only "random_samp10"" returns a mean. This makes sense because it is the only data column with numerica data. "more_than_2", "vec_cities", and "vec_continent", all contain string data and thus a mean does not make sense with this data. 

##Changing variables to numeric
```{r changing_numeric, results = FALSE}
help("as.numeric")
example("as.numeric")
as.numeric(problem1_df[["more_than_2"]])
as.numeric(problem1_df[["vec_cities"]])
as.numeric(problem1_df[["vec_continent"]])
```
Using as.numeric on the logical("more_than_2") and character ("vec_cities") variables changes all the data to "NA" and warns us that it created these "NA" by coercion. It does not know how to interpret these strings as numbers. 

Using as.numeric on the factor ("vec_continent") varaible assigns numbers to each category, counting up using alphabetical order. 
